Laplacian envelope

### 2021-09-29 Notes/laplace.tex
started scaling argument

\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
%================================================================================
%
% \usepackage{adjustbox}
% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
% %
% \usepackage{paralist}
% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
% %for nested lists, \ref only produces innermost item:  (hacky)
% \makeatletter % for internal macros with @
% \renewcommand\p@enumii{}
% \makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%prob distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diag first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
\newcommand{\dgnm}{\zeta}
\newcommand{\dgn}{\boldsymbol{\dgnm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
% %avoid double superscripts
% \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

@@

The area under the SNR curve is:
%
\begin{equation}\label{eq:areadef}
  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) 
      = A(0) = \frac{\sqrt{N}}{r} \CA = \sqrt{N} \sum_a \CI_a \tau_a.
\end{equation}
%
To maximise this quantity, consider the Lagrangian
%
\begin{equation}\label{eq:arealagrangian}
\begin{aligned}
  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(0) - \etwm_j(0))
        +\sum_{\nu i j} \prn{ \mu^\nu_{ij} \MMdm^\nu_{ij} }
        &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
\end{aligned}
\end{equation}
%
where \(\Delta p\) is the constant value we are holding \(\eq\w\) at.
The Karush-Kuhn-Tucker necessary conditions for a maximum are
%
\begin{equation}\label{eq:KTarea}
  \pdiff{\CL}{\MMdm^\nu_{ij}} = 0, \quad
  \mu^\nu_{ij} \geq 0, \quad
  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
  \pdiff{\CL}{\lambda} = 0.
\end{equation}
%
From \cref{eq:lAdiff}, the derivatives for \(m \neq n\) are
%
\begin{equation}\label{eq:areaderivs}
\begin{aligned}
  \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
    r f^\nu \eqm_m \prn{ \brk{\thbm_m(0) - \thbm_n(0)}
     + \prn{c_m(0) + (-1)^\nu} \brk{\etwm_m(0) - \etwm_n(0)} } \\
     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
     + \lambda r f^\nu \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
    = 0.
\end{aligned}
\end{equation}
%
See \cref{eq:lareacoeffs} for the definitions of these quantities.

The area and \(\Delta p\) are invariant under the scale transformation %\cref{eq:scale}
%
\begin{equation}\label{eq:scale}
  \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I .
\end{equation}
%
It follows that
%
\begin{equation*}
  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} 
    = - \sum_{\nu, m} \mu^\nu_{mm} = 0 .
\end{equation*}
%
As the KKT multipliers are nonnegative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.

@@

Consider
%
\begin{equation}\label{eq:shiftqderiv}
  2r\pdiff{\CL}{\encm_{mn}} =
  \frac{1}{f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} - \frac{1}{f\pot} \pdiff{\CL}{\MMdm\dep_{mn}}
   = \mu\pot_{mn} - \mu\dep_{mn} + 2\eqm_m (\etwm_m - \etwm_n)
   = 0.
\end{equation}
%
This corresponds to the shift
%
\begin{equation}\label{eq:shiftq}
  \MMdm\pot_{ij} \to \MMdm\pot_{ij} + \frac{\epsilon_{ij}}{f\pot},
  \qquad
  \MMdm\dep_{ij} \to \MMdm\dep_{ij} - \frac{\epsilon_{ij}}{f\dep},
  \qquad
  \sum_j \epsilon_{ij} = 0,
\end{equation}
%
which leaves $\frg$ unchanged, and therefore $\eq$, $\fptb$ and $\etw$ as well.

Assume $\etwm_m > \etwm_n$. Then
%
\begin{equation}\label{eq:lowertriangular}
 \mu\pot_{mn} - \mu\dep_{mn} < 0
 \qquad\implies\qquad
 \mu\dep_{mn} > 0
 \qquad\implies\qquad
 \MMdm\dep_{mn} = 0.
\end{equation}
%
Similarly, if $\etwm_m < \etwm_n$, then
%
\begin{equation}\label{eq:uppertriangular}
 \mu\pot_{mn} - \mu\dep_{mn} > 0
 \qquad\implies\qquad
 \mu\pot_{mn} >0
 \qquad\implies\qquad
 \MMdm\pot_{mn}=0.
\end{equation}
%
Thus, if we arrange the states in order of decreasing $\etw$, $\MMd\pot$ is upper-triangular and $\MMd\dep$ is lower triangular.

@@

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
%
\begin{equation}\label{eq:sareacoeffincrprob}
  \Delta^\nu_{mn} \eqm_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:areacoeffincrfinite}
  \MMd^\nu \to \MMd^\nu + \Eq\inv \boldsymbol{\epsilon},
  \qquad
  \begin{aligned}
  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
  \\
  \boldsymbol{\epsilon} \onev &= 0.
  \end{aligned}
\end{equation}
%
But more interestingly:
%
\begin{equation}\label{eq:areacoeffincr}
    \Delta^\nu_{mn}\CL =
      \frac{\mu^\nu_{mn}}{\eqm_m} + \frac{\mu^\nu_{nm}}{\eqm_n}
      + r \prn{c_m - c_n}\prn{\etwm_m - \etwm_n} = 0,
\end{equation}
%
where $c_k$ were defined in \cref{eq:lareacoeffs}, with \(s = 0\).

Using the non-negativity of the Kuhn-Tucker multipliers, $\mu^\nu_{ij}$, \cref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\etwm_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\etwm_k$ are degenerate, we can choose their order to ensure this).

Note that, according to \cref{sec:triangular}, either $\MMdm\pot_{mn}$ or $\MMdm\dep_{mn}$ will be zero at the maximum, therefore we can expect one of the $\mu^\nu_{mn}$ to be non-zero.
This would rule out degeneracy of the $c_k$ or $\etwm_k$.
Looking at \cref{eq:shiftqderiv} closely, the only way both $\mu^\nu_{mn}$ could be zero is if $\etwm_m = \etwm_n$ or $\eqm_m = \eqm_n = 0$.

@@

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
%
\begin{equation}\label{eq:shortcutprob}
  \widetilde{\Delta}^\nu_{g,m} \eqm_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:shortcutfinite}
  \MMd^\nu \to \MMd^\nu + \Eq\inv \boldsymbol{\epsilon}^{\nu mn},
  \qquad
  \begin{alignedat}[t]{3}
    &\prn{{\epsilon}^{\nu mn}}_{mn}
      &&= -\epsilon,\\
    &\prn{{\epsilon}^{\nu mn}}_{k,k\pm1}
      &&= \epsilon
      &\;&\forall\, k \in [m,n),\\
    &\prn{{\epsilon}^{\nu mn)}}_{kk}
      &&= -\epsilon
      &&\forall\, k \in (m,n).
  \end{alignedat}
\end{equation}
%
But more interestingly for our purposes:
%
\begin{equation}\label{eq:shortcutarea}
    \widetilde{\Delta}^\nu_{mn}\CL =
      \brk{ \sum_{k=m}^{n\mp1} \frac{\mu^\nu_{k,k\pm1}}{\eqm_k}
      - \frac{\mu^\nu_{m,n}}{\eqm_m}}
      + rf^\nu \sum_{k=m}^{n\mp1} \prn{\etwm_{k} - \etwm_{k\pm1}} \prn{c_k - c_m},
\end{equation}
%

If we put the states in order of decreasing $\etwm_k$, the results of the \cref{sec:areacoeffincr} tell us that the $c_k$ are non-decreasing.
This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu^\nu_{mn} > 0$, which in turn implies that $\MMdm^\nu_{mn}=0$.
This would tell us that the process with the maximal area has to have a serial topology.

@@

@@

So we see that all four inequalities cannot be simultaneously saturated, and the bound must be loose.

@@

Consider the Lagrangian
%
\begin{equation}\label{eq:envlagrangian}
  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
\end{equation}
%
The Karush-Kuhn-Tucker necessary conditions for a maximum are
%
\begin{equation}\label{eq:envKTcond}
  \pdiff{\CL}{\CI_a} = 0, \qquad
  \pdiff{\CL}{\tau_a} = 0, \qquad
  \mu_\alpha \geq 0, \qquad
  \pdiff{\CL}{\mu_\alpha} \geq 0, \qquad
  \mu_\alpha\pdiff{\CL}{\mu_\alpha} = 0.
\end{equation}
%
These are solved by
%
\begin{equation}\label{eq:envsol}
\begin{gathered}
  \CI_1 = 1, \qquad
  \CI_{a>1} = 0, \qquad
  \tau_1 = M-1, \\
  \mu_\CI = \frac{(M-1)^2}{[r\tau + (M-1)]^2}, \qquad
  \mu_\CA = \frac{r\tau}{[r\tau+(M-1)]^2}. \qquad
\end{gathered}
\end{equation}
%
This leads to the envelope
%
\begin{equation}\label{eq:env}
  \snrb(\tau) \leq \frac{\sqrt{N}(M-1)}{r\tau + (M-1)}.
\end{equation}
%
This is the envelope (``memory frontier'') that we can prove.
Finding more constraints would lower it.
Clearly there have to be more constraints, as \eqref{eq:envsol} indicates that this would be achieved by the same model at all times.
This model would have to saturate both the initial SNR bound \eqref{eq:initbnd} and the area bound \eqref{eq:areabnd}, whereas we saw that they were achieved by different models, which perform well at short/long timescales respectively.

@@

If we know that the optimal model for any timescale satisfies detailed balance, we have one more constraint, \cref{eq:sqrtbound}.
Then the Lagrangian reads
%
\begin{equation}\label{eq:envDBlagrangian}
  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} 
      + \sum_a \mu_a \prn{\sqrt{2} - \CI_a \sqrt{\tau_a}} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
\end{equation}
%
The solution to \cref{eq:envKTcond} takes different forms for different ranges of \(\tau\) and \(M\),
as described in \cref{tab:envDB}.
The resulting envelope is (for \(M > 3\))
%
\begin{equation}\label{eq:envDB}
  \snrb(\tau) \leq 
  \left\{ 
  \begin{alignedat}{3}
    &\frac{2 \sqrt{N}}{2 + r \tau}
      , &
      0 &\leq r \tau &&\leq 2
      , \\
    &\sqrt{\frac{N}{2 r \tau}}
      , &
      2 &< r \tau &&< \frac{(M-1)^2}{2}
      , \\
    &\frac{2 \sqrt{N} (M-1)}{(M-1)^2 + 2r \tau}
      , &\qquad
      \frac{(M-1)^2}{2} &\leq r \tau 
      .
  \end{alignedat}
  \right.
\end{equation}
%
This envelope is shown in \cref{fig:envproven}(\ref{fig:provendetbal}).
For \(M \leq 3\), \cref{eq:env} still applies.

@@

The Lagrangian we should consider is
%
\begin{equation}\label{eq:lshiftlagrangian}
\begin{aligned}
  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
        +\sum_{\nu i j} \mu^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
        &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
\end{aligned}
\end{equation}
%
where all of the first-passage times are evaluated at \(s = 0\).

>>>>

\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
%================================================================================
%
% \usepackage{adjustbox}
% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
% %
% \usepackage{paralist}
% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
% %for nested lists, \ref only produces innermost item:  (hacky)
% \makeatletter % for internal macros with @
% \renewcommand\p@enumii{}
% \makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%prob distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diag first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
\newcommand{\dgnm}{\zeta}
\newcommand{\dgn}{\boldsymbol{\dgnm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%KKT multipliers
\newcommand{\kktm}{\mu}
\newcommand{\kkt}{\boldsymbol{\kktm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
\newcommand{\cnst}{\gamma}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
% %avoid double superscripts
% \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

@@

The area under the SNR curve is:
%
\begin{equation}\label{eq:areadef}
  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) 
      = A(0) = \frac{\sqrt{N}}{r} \CA = \sqrt{N} \sum_a \CI_a \tau_a.
\end{equation}
%
To maximise this quantity, consider the Lagrangian
%
\begin{equation}\label{eq:arealagrangian}
\begin{aligned}
  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(0) - \etwm_j(0))
        +\sum_{\nu i j} \prn{ \kktm^\nu_{ij} \MMdm^\nu_{ij} }
        &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
\end{aligned}
\end{equation}
%
where \(\Delta p\) is the constant value we are holding \(\eq\w\) at.
The Karush-Kuhn-Tucker necessary conditions for a maximum are
%
\begin{equation}\label{eq:KTarea}
  \pdiff{\CL}{\MMdm^\nu_{ij}} = 0, \quad
  \kktm^\nu_{ij} \geq 0, \quad
  \pdiff{\CL}{\kktm^\nu_{ij}} \geq 0, \quad
  \kktm^\nu_{ij}\pdiff{\CL}{\kktm^\nu_{ij}} = 0, \quad
  \pdiff{\CL}{\lambda} = 0.
\end{equation}
%
From \cref{eq:lAdiff}, the derivatives for \(m \neq n\) are
%
\begin{equation}\label{eq:areaderivs}
\begin{aligned}
  \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
    r f^\nu \eqm_m \prn{ \brk{\thbm_m(0) - \thbm_n(0)}
     + \prn{c_m(0) + (-1)^\nu} \brk{\etwm_m(0) - \etwm_n(0)} } \\
     &+ \kktm^\nu_{mn} -  \kktm^\nu_{mm}
     + \lambda r f^\nu \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
    = 0.
\end{aligned}
\end{equation}
%
See \cref{eq:lareacoeffs} for the definitions of these quantities.

The area and \(\Delta p\) are invariant under the scale transformation %\cref{eq:scale}
%
\begin{equation}\label{eq:scale}
  \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I .
\end{equation}
%
It follows that
%
\begin{equation*}
  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} 
    = - \sum_{\nu, m} \kktm^\nu_{mm} = 0 .
\end{equation*}
%
As the KKT multipliers are nonnegative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.

@@

Consider
%
\begin{equation}\label{eq:shiftqderiv}
  2r\pdiff{\CL}{\encm_{mn}} =
  \frac{1}{f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} - \frac{1}{f\pot} \pdiff{\CL}{\MMdm\dep_{mn}}
   = \kktm\pot_{mn} - \kktm\dep_{mn} + 2\eqm_m (\etwm_m - \etwm_n)
   = 0.
\end{equation}
%
This corresponds to the shift
%
\begin{equation}\label{eq:shiftq}
  \MMdm\pot_{ij} \to \MMdm\pot_{ij} + \frac{\epsilon_{ij}}{f\pot},
  \qquad
  \MMdm\dep_{ij} \to \MMdm\dep_{ij} - \frac{\epsilon_{ij}}{f\dep},
  \qquad
  \sum_j \epsilon_{ij} = 0,
\end{equation}
%
which leaves $\frg$ unchanged, and therefore $\eq$, $\fptb$ and $\etw$ as well.

Assume $\etwm_m > \etwm_n$. Then
%
\begin{equation}\label{eq:lowertriangular}
 \kktm\pot_{mn} - \kktm\dep_{mn} < 0
 \qquad\implies\qquad
 \kktm\dep_{mn} > 0
 \qquad\implies\qquad
 \MMdm\dep_{mn} = 0.
\end{equation}
%
Similarly, if $\etwm_m < \etwm_n$, then
%
\begin{equation}\label{eq:uppertriangular}
 \kktm\pot_{mn} - \kktm\dep_{mn} > 0
 \qquad\implies\qquad
 \kktm\pot_{mn} >0
 \qquad\implies\qquad
 \MMdm\pot_{mn}=0.
\end{equation}
%
Thus, if we arrange the states in order of decreasing $\etw$, $\MMd\pot$ is upper-triangular and $\MMd\dep$ is lower triangular.

@@

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
%
\begin{equation}\label{eq:sareacoeffincrprob}
  \Delta^\nu_{mn} \eqm_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:areacoeffincrfinite}
  \MMd^\nu \to \MMd^\nu + \Eq\inv \boldsymbol{\epsilon},
  \qquad
  \begin{aligned}
  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
  \\
  \boldsymbol{\epsilon} \onev &= 0.
  \end{aligned}
\end{equation}
%
But more interestingly:
%
\begin{equation}\label{eq:areacoeffincr}
    \Delta^\nu_{mn}\CL =
      \frac{\kktm^\nu_{mn}}{\eqm_m} + \frac{\kktm^\nu_{nm}}{\eqm_n}
      + r \prn{c_m - c_n}\prn{\etwm_m - \etwm_n} = 0,
\end{equation}
%
where $c_k$ were defined in \cref{eq:lareacoeffs}, with \(s = 0\).

Using the non-negativity of the Kuhn-Tucker multipliers, $\kktm^\nu_{ij}$, \cref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\etwm_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\etwm_k$ are degenerate, we can choose their order to ensure this).

Note that, according to \cref{sec:triangular}, either $\MMdm\pot_{mn}$ or $\MMdm\dep_{mn}$ will be zero at the maximum, 
therefore we can expect one of the $\kktm^\nu_{mn}$ to be non-zero.
This would rule out degeneracy of the $c_k$ or $\etwm_k$.
Looking at \cref{eq:shiftqderiv} closely, the only way both $\kktm^\nu_{mn}$ could be zero is if $\etwm_m = \etwm_n$ or $\eqm_m = \eqm_n = 0$.

@@

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
%
\begin{equation}\label{eq:shortcutprob}
  \widetilde{\Delta}^\nu_{g,m} \eqm_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:shortcutfinite}
  \MMd^\nu \to \MMd^\nu + \Eq\inv \boldsymbol{\epsilon}^{\nu mn},
  \qquad
  \begin{alignedat}[t]{3}
    &\prn{{\epsilon}^{\nu mn}}_{mn}
      &&= -\epsilon,\\
    &\prn{{\epsilon}^{\nu mn}}_{k,k\pm1}
      &&= \epsilon
      &\;&\forall\, k \in [m,n),\\
    &\prn{{\epsilon}^{\nu mn)}}_{kk}
      &&= -\epsilon
      &&\forall\, k \in (m,n).
  \end{alignedat}
\end{equation}
%
But more interestingly for our purposes:
%
\begin{equation}\label{eq:shortcutarea}
    \widetilde{\Delta}^\nu_{mn}\CL =
      \brk{ \sum_{k=m}^{n\mp1} \frac{\kktm^\nu_{k,k\pm1}}{\eqm_k}
      - \frac{\kktm^\nu_{m,n}}{\eqm_m}}
      + rf^\nu \sum_{k=m}^{n\mp1} \prn{\etwm_{k} - \etwm_{k\pm1}} \prn{c_k - c_m},
\end{equation}
%

If we put the states in order of decreasing $\etwm_k$, the results of the \cref{sec:areacoeffincr} tell us that the $c_k$ are non-decreasing.
This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\kktm^\nu_{mn} > 0$, which in turn implies that $\MMdm^\nu_{mn}=0$.
This would tell us that the process with the maximal area has to have a serial topology.

@@

Note that in the case of a model that is not of the serial topology, but is lumpable to one, 
it will satisfy \cref{eq:areabnd} with a lower value of \(M\).
Therefore the maximal area will come from a model with the full serial topology.

@@

So we see that all four inequalities cannot be simultaneously saturated, and the bound must be loose.
Because we think that \cref{eq:sqrtbound} can be improved, we will rewrite it as
%
\begin{equation}\label{eq:gammabnd}
  \CI_a \sqrt{r\tau_a} \leq \sqrt{\cnst}.
\end{equation}
% 
We have shown that the smallest \(\cnst\) for which \cref{eq:gammabnd} is true is less than 2.
Considering the two-state model with deterministic transitions, we see that \(\cnst\) is at least 1.
Later, looking at uniform serial models in \cref{sec:heuristicenv} we will see near \cref{eq:henvapprox} that \(\gamma\) is at least 1.18.

@@

Consider the Lagrangian
%
\begin{equation}\label{eq:envlagrangian}
  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \kktm_\CI \prn{1 - \sum_a \CI_a} + \kktm_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
\end{equation}
%
The Karush-Kuhn-Tucker necessary conditions for a maximum are
%
\begin{equation}\label{eq:envKTcond}
  \pdiff{\CL}{\CI_a} = 0, \qquad
  \pdiff{\CL}{\tau_a} = 0, \qquad
  \kktm_\alpha \geq 0, \qquad
  \pdiff{\CL}{\kktm_\alpha} \geq 0, \qquad
  \kktm_\alpha\pdiff{\CL}{\kktm_\alpha} = 0.
\end{equation}
%
These are solved by
%
\begin{equation}\label{eq:envsol}
\begin{gathered}
  \CI_1 = 1, \qquad
  \CI_{a>1} = 0, \qquad
  \tau_1 = M-1, \\
  \kktm_\CI = \frac{(M-1)^2}{[r\tau + (M-1)]^2}, \qquad
  \kktm_\CA = \frac{r\tau}{[r\tau+(M-1)]^2}. \qquad
\end{gathered}
\end{equation}
%
This leads to the envelope
%
\begin{equation}\label{eq:env}
  \snrb(\tau) \leq \frac{\sqrt{N}(M-1)}{r\tau + (M-1)}.
\end{equation}
%
This is the envelope (``memory frontier'') that we can prove.
Finding more constraints would lower it.
Clearly there have to be more constraints, as \eqref{eq:envsol} indicates that this would be achieved by the same model at all times.
This model would have to saturate both the initial SNR bound \eqref{eq:initbnd} and the area bound \eqref{eq:areabnd}, whereas we saw that they were achieved by different models, which perform well at short/long timescales respectively.

@@

If we know that the optimal model for any timescale satisfies detailed balance, we have one more constraint, \cref{eq:sqrtbound}.
Then the Lagrangian reads
%
\begin{equation}\label{eq:envDBlagrangian}
  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \kktm_\CI \prn{1 - \sum_a \CI_a} 
      + \sum_a \kktm_a \prn{\sqrt{\cnst} - \CI_a \sqrt{\tau_a}} + \kktm_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
\end{equation}
%
The solution to \cref{eq:envKTcond} takes different forms for different ranges of \(\tau\) and \(M\),
as described in \cref{tab:envDB}.
The resulting envelope is (for \(M > 3\))
%
\begin{equation}\label{eq:envDB}
  \snrb(\tau) \leq 
  \left\{ 
  \begin{alignedat}{3}
    &\frac{\sqrt{N} \cnst}{\cnst + r \tau}
      , &
      0 &\leq r \tau &&\leq \cnst
      , \\
    &\sqrt{\frac{N\cnst}{4 r \tau}}
      , &
      \cnst &< r \tau &&< \frac{(M-1)^2}{\cnst}
      , \\
    &\frac{\sqrt{N} \cnst (M-1)}{(M-1)^2 + \cnst r \tau}
      , &\qquad
      \frac{(M-1)^2}{\cnst} &\leq r \tau 
      .
  \end{alignedat}
  \right.
\end{equation}
%
This envelope is shown in \cref{fig:envproven}(\ref{fig:provendetbal}).
For \(M \leq 3\), \cref{eq:env} still applies.

@@

The Lagrangian we should consider is
%
\begin{equation}\label{eq:lshiftlagrangian}
\begin{aligned}
  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
        +\sum_{\nu i j} \kktm^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
        &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
\end{aligned}
\end{equation}
%
where all of the first-passage times are evaluated at \(s = 0\).
